FraCaS: 80 percent complete

### 2019-02-26 paper2018/paper.tex
various minor revisions

\begin{abstract}
  TODO
\end{abstract}
\section{Introduction}
Natural Language Inference (NLI) is the task of determining of whether an NL hypothesis H follows from an NL premise(s) P. NLI has received a lot of attention in the Computational Semantics literature and has been approached using a variety of techniques, ranging from logical approaches  \citet{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017} all the way to the recent Deep Learning (DL) models for NLI. The latter approaches, following the general trend in NLP, have been dominating NLI and indeed a number of impressive results have been produced \citet{kim:2018,radford:2018,liu:2019}.\footnote{These are the three systems with the best results on SNLI in increasing order.} State-of-the-art DL systems achieve an accuracy of around 0.9 when tested on suitable datasets. However, the datasets that are used are assuming a definition of inference that can be thought to be `looser' or less precise compared to the definition assumed in platforms that used to test logical approaches. Consider the following two examples, one from the SNLI dataset, predominatly used to test DL approaches and which has a state-of-the-art performace of 0.911 \citet{lui:2019'}, and the FraCaS test suite, the dataset \textit{par excellence} to test logical approaches.

\begin{lingex}
\item An SNLI example \\
\textbf{P}  A  man selling donuts to a customer during a world
exhibition event held in the city of Angeles. \\
 \textbf{H}  A woman drinks her coffee in a small cafe.\\
\textbf{Label: Contradiction}
\end{lingex}

\begin{lingex}
\item An example from FraCaS\\

\textbf{P1}] A Scandinavian won the Nobel Prize.\\
\textbf{P2}	Every Swede is  Scandinavian.\\
\textbf{Q}  Did a Swede win the Nobel prize? \\
\textbf{H} A Swede won the Nobel prize.\\
\textbf{Label: Yes} [FraCaS 173]
\end{lingex}

In the SNLI example, one has to make a number of non-trivial assumptions in order to arrive at a contradiction:  a)  the two situations described have to be taken to refer to  the same situation and, thus, decide  whether one description contradicts the other, b) the indefinite article in the premise has to be identified with the indefinite article in the hypothesis, i.e.  \textit{man} contradicting \textit{woman} (a person cannot be a man and a woman at the same time). This is part of the reasoning humans perform, but it is not the only one. More precise, logical reasoning can be also performed. Such an example is the FraCaS example, as well as very fine-grained cases like the Antecedent Contained Deletion  ellipsis example below:

@@

For this reason, and despite the dominance of DL approaches in pretty much all NLP tasks, logical approaches have been continuing to be developed and evaluated on datasets like the FraCaS test suite and the SICK dataset \cite{marelli:2014}.  \citet{bernardy:2017}  define  a correspondence between abstract syntax parse trees of the FraCas examples, parsed using the Grammatical Framework (GF, \citet{Ranta:GF}), and modern type-theoretic semantics that are output in the Coq proof assistant (the FraCoq system).  The accuracy is 0.85 for 5 sections of the FraCaS test suite. The \textsc{Langpro} system presented by \citet{abzianidze:2017} is based on a Natural Logic tableau theorem prover. It achieves an accuracy of .82 on the SICK dataset.

@@

\begin{itemize}
 \item  A two-phase system (rather than one-phase as in  \citet{bernardy_type_2017}): a) using Haskell as a converter between Grammatical Framework trees and Coq types, b) using various type combinators to reach the desired NL semantics. 
  \item  The first run of the Anaphora and Ellipsis sections of the FraCaS (based on work by \citet{bernardy_jolli})
 \item  More coverage than any previous logical approach
  \item  State-of-the-art results in terms of accuracy compared to all previous approaches, for every section of the FraCaS
  \item  Handing of multiple readings
  \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
  \item  New handling of prepositions and adverbs compared to the earlier system in  \citet{bernardy_type_2017}, as well as new handling of definitions, enabled by the dynamic semantics system. 
  \item  Improvement over the previous tre
atment of Plurals/Quantifiers
 \item  Big improvement on the comparatives section  enabled by handling  comparatives using a threshold (BUT one different to the one used for adjectives). Again, the dynamic semantics system is crucial here.

@@

\subsection{GF}
In GF, abstract syntax is comprised of: a) a number of syntactic
categories, and b) a number of syntactic construction functions. The latter
provide the means to compose basic syntactic categories into more complex
ones.  For example, consider the constructor: $AdjCN : AP → CN → CN$. This  expresses
that one can append an adjectival phrase to a common noun and obtain
a new common noun. Furthermore,  GF is equipped  with a library of mappings from abstract
syntax to the concrete s usyntax of varionatural languages. These mappings can be inverted by GF, thus offering
parsing from natural text into abstract syntax. However, in this project we skip the
parsing phase and use the parse trees constructed by \citet{Ljunglof:2012},
thereby avoiding any syntactic ambiguity. \textbf{Do you want to further comment here?}

\subsection{Coq}
Coq is an  interactive theorem prover (proof assistant)  based on the calculus of inductive
constructions (CiC), i.e.  a lambda calculus with dependent
types. Coq is arguably one of the leading proof assistants. It has
famously been instrumental in formalising a proof of the four-color
theorem \citep{Gonthier:2008}, a proof of the odd order theorem \citep{Gonthier:2013}, as well as developing CompCert, a formally verified compiler for C \citep{Leroy}. Coq is a very powerful reasoning engine that makes it fit for the task of NLI, when the latter is formalized as theorem proving task. Its underlying higher order logic is furthermore very expressive to express fine-grained logical semantics. Furthermore, Coq supports a number of constructions that we will be using notably dependent typing and subtyping. These features are used in order to take care of specific NL phenomena (e.g. subtyping is used in defining different adjectival inference classes).

\subsection{Dynamic Monadic Semantics}
Dynamic Monadic Semantics have been proven to be an effective way of dealing with anaphora and ellipsis. There are a number of  approaches using monads or other similar ideas (e.g. continuations as in \cite{de2006}) for anaphora and ellipsis  \citet{Shan:2002,unger:2011,Barker:04,de2016,charlow:2017}. In this paper, we follow the approach  as described in \citet{bernardy_jolli}. More details on this will be given in the next section, where the actual system is presented.

@@

Our system consists of two main parts.
\begin{enumerate}
\item A converter from syntax trees to types. The syntax trees follow
  the GF formalism, and the types follow the Coq formalism. The
  converter itself is a Haskell Program. It implements a dynamic
  semantics. It comprises the bigger part of our system.
\item A number of type-theoretical combinators. These take care of the
  natural semantics which have no influence on the dynamical
  aspects. This includes for example the treatment of adjectives
  (intersective, subsective, etc.) and adverbs (veridical or not).
\end{enumerate}
This architecture is represented schematically in Figure \ref{fig:overview}.

@@

All underlying systems (GF, Haskell, Coq) are based on lambda calculi
with types. We ensure that each translation preserve typing, locally:
\begin{enumerate}
\item Every GF syntactic category $C$ is mapped to a type noted $⟦C⟧$.
\item GF Functional types are mapped compositionally : $⟦A → B⟧ = ⟦A⟧ → ⟦B⟧$
\item Every GF syntactic construction function $f : X$ is mapped to a function $⟦f⟧$ such that $⟦f⟧ : ⟦X⟧$.
\item GF function applications are mapped compositionally: $⟦t(u)⟧ = ⟦t⟧ (⟦u⟧)$.
\end{enumerate}
Because all system embed the simply-typed lambda calculus, ensuring
type-preservation locally means that types are preserved globally.
Therefor, we are certain that every GF syntax tree can be mapped to
Haskell, and eventuall Coq, without error.

@@

\paragraph{Adjectives}
In our system, we interpret gradable adjectives using a pair of a
measure $m : objects → Z$ and and a threshold $τ : Z$, where $Z$ is
treated as an abstract ordered ring by Coq. (This structure has no
dynamic aspect in our model, and thus is entirely handled within Coq.)
For subsective adjectives, $τ$ will additionally depend on the class
of the object in question. This structure has the benefit that
opposite adjectives can be easily represented (measures are opposites
$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 > 0$). Formalization aside, this idea is reminiscent of degree-based approaches to gradable adjectives like \citet{Cresswell:1976,Heim:1985,Kennedy:2007}.   Additionally,
adjectival phrases, as present in the FraCaS suite, are interpreted as
linear inequations in $Z$. Solving systems of such inequations is
decidable. Indeed, the tactic that Coq offers for this purpose can
solve all such problems in the FraCaS suite, automatically.

@@

\paragraph{error analysis}
Our system fails to classify correctly 28 cases out of 259.  We give
here a summary of the missing features which are responsible for the
biggest number of failures. The one biggest source of error is an
incomplete handling of group readings. (FraCaS
013, 014, 046, 084, 111, 124, 126, 127, 137, 171, 172, 191, 193, 195, 243, 250, 333, 346). These
are cases where a syntactic conjunction of individuals is treated as a
semantic group, or where precise counting of the members of a group is
necessary.  Other problematic cases are where a definite plural does
not have a universal reading (091,094,095) --- these require an
analysis at the pragmatic level, which our system lacks. Additionally,
neither measure phrases (242) nor attributive comparatives (244,245)
are handled.

@@

\bibliography{PaperTools/bibtex/jp.bib,phorae.bib,IWCSbib.bib}

>>>>

\begin{abstract}
  TODO
\end{abstract}
\section{Introduction}
Natural Language Inference (NLI) is the task of determining of whether
an NL hypothesis H follows from an NL premise(s) P. NLI has received a
lot of attention in the Computational Semantics literature and has
been approached using a variety of techniques, ranging from logical
approaches
\citep{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017},
all the way to the recent Deep Learning (DL) models for NLI. The
latter approaches, following a general trend in NLP, have been
dominating NLI and a number of impressive results have been
produced \citep{kim:2018,radford:2018,liu:2019}.\footnote{These are
  the three systems with the best results on SNLI in increasing
  order at the time of writing.}  State-of-the-art DL systems achieve an accuracy of around
0.9 when tested on suitable datasets. However, the datasets that are
used are assuming a definition of inference that can be thought to be
`looser' or less precise compared to the definition assumed in
platforms based in logical approaches. Consider the following
two examples, one from the SNLI dataset, predominatly used to test DL
approaches and which has a state-of-the-art performace of 0.911
\citet{lui:2019'}, and the FraCaS test suite, the dataset \textit{par
  excellence} to test logical approaches.
%
The following two examples illustrate the issue.
\begin{lingex}
\item % An SNLI example \\
\textbf{P}  A  man selling donuts to a customer during a world
exhibition event held in the city of Angeles. \\
 \textbf{H}  A woman drinks her coffee in a small cafe.\\
\textbf{Label: Contradiction} [SNLI]
\item % An example from FraCaS\\
\textbf{P1}] A Scandinavian won the Nobel Prize.\\
\textbf{P2}	Every Swede is  Scandinavian.\\
\textbf{Q}  Did a Swede win the Nobel prize? \\
\textbf{H} A Swede won the Nobel prize.\\
\textbf{Label: Yes} [FraCaS 173]
\end{lingex}
%
In the SNLI example, one has to make a number of non-trivial
assumptions in order to arrive at a contradiction: a) the two
situations described have to be taken to refer to the same situation
and, thus, decide whether one description contradicts the other, b)
the indefinite article in the premise has to be identified with the
indefinite article in the hypothesis. (Additionally considering that a
person cannot be a man selling donuts and a woman drinking coffee at
the same time.) While this can be part of the reasoning humans
perform, it is not the only possibility. More precise, logical
reasoning can be also performed. Such an example is the FraCaS
example, as well as very fine-grained cases like the Antecedent
Contained Deletion ellipsis example below:

@@

For this reason, and despite the dominance of DL approaches in pretty
much all NLP tasks, logical approaches continue to be
developed and evaluated on datasets like the FraCaS test suite and the
SICK dataset \cite{marelli:2014}.  \citet{bernardy:2017} define a
correspondence between abstract syntax parse trees of the FraCas
examples, parsed using the Grammatical Framework (GF,
\citet{Ranta:GF}), and modern type-theoretic semantics that are output
in the Coq proof assistant (the FraCoq system).  The accuracy is 0.85
for 5 sections of the FraCaS test suite. The \textsc{Langpro} system
presented by \citet{abzianidze:2017} is based on a Natural Logic
tableau theorem prover. It achieves an accuracy of .82 on the SICK
dataset.

@@

\begin{itemize}
\item A two-phase system (rather than a one-phase as described by
  \citet{bernardy_type_2017}): a) using Haskell as a converter between
  Grammatical Framework trees and Coq types, b) using various type
  combinators to reach the desired NL semantics.
\item The first run of the Anaphora and Ellipsis sections of the
  FraCaS (based on work by \citet{bernardy_jolli})
 \item  More coverage than any previous logical approach
 \item State-of-the-art results in terms of accuracy compared to all
   previous approaches, for every section of the FraCaS
  \item Handing of multiple readings
  \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
  \item New handling of prepositions and adverbs compared to the
    earlier system in \citet{bernardy_type_2017}, as well as new
    handling of definitions, enabled by the dynamic semantics system.
  \item Improvement over the previous treatment of Plurals/Quantifiers
  \item Big improvement on the comparatives section enabled by
    handling comparatives using a threshold (BUT one different to the
    one used for adjectives). Again, the dynamic semantics system is
    crucial here.

@@

\subsection{GF}
In GF, abstract syntax is comprised of: a) a number of syntactic
categories, and b) a number of syntactic construction functions. The latter
provide the means to compose basic syntactic categories into more complex
ones.  For example, consider the constructor: $AdjCN : AP → CN → CN$. This  expresses
that one can append an adjectival phrase to a common noun and obtain
a new common noun. Furthermore,  GF is equipped  with a library of mappings from abstract
syntax to the concrete s usyntax of varionatural languages. These mappings can be inverted by GF, thus offering
parsing from natural text into abstract syntax. However, in this project we skip the
parsing phase and use the parse trees constructed by \citet{Ljunglof:2012},
thereby avoiding any syntactic ambiguity. \todo{Do you want to further comment here?}

\subsection{Coq}
Coq is an interactive theorem prover (proof assistant) based on the
calculus of inductive constructions (CiC), i.e.  a lambda calculus
with dependent types. Coq is arguably one of the leading proof
assistants. It has famously been instrumental in formalising a proof
of the four-color theorem \citep{Gonthier:2008}, a proof of the odd
order theorem \citep{Gonthier:2013}, as well as developing CompCert, a
formally verified compiler for C \citep{Leroy}. Coq is a very powerful
reasoning engine that makes it fit for the task of NLI, when the
latter is formalized as theorem proving task. Its underlying higher
order logic is furthermore powerful enough to express fine-grained
logical semantics. Furthermore, Coq supports a number of constructions
that we will be using notably dependent typing and subtyping. These
features allow us to take care of specific NL phenomena
(e.g. subtyping is used in defining different adjectival inference
classes).

\subsection{Dynamic Monadic Semantics}
Dynamic Monadic Semantics have been proven to be an effective way of
dealing with anaphora and ellipsis. There are a number of approaches
using monads or other equivalent constructions (e.g. continuations as in the work of
\citet{de2006}) for anaphora and ellipsis
\citet{Shan:2002,unger:2011,Barker:04,de2016,charlow:2017}. In this
paper, we follow the approach described by
\citet{bernardy_jolli}. More details on this will be given in the next
section, where the actual system is presented.

@@

Our system consists of two main parts.
\begin{enumerate}
\item A converter from syntax trees to types. The syntax trees follow
  the GF formalism, and the types follow the Coq formalism. The
  converter itself is a Haskell Program, which implements a dynamic
  semantics and comprises the bulk of our system.
\item A number of type-theoretical combinators. These take care of the
  natural semantics which have no influence on the dynamical
  aspects. Such aspects include for example the treatment of adjectives
  (intersective, subsective, etc.) and adverbs (veridical or not).
\end{enumerate}
The architecture is represented schematically in Figure \ref{fig:overview}.

@@

All the underlying systems (GF, Haskell, Coq) are based on lambda
calculi with types. We take advantage of typing, ensuring that each
translation preserve typing, locally:
\begin{enumerate}
\item Every GF syntactic category $C$ is mapped to a type noted $⟦C⟧$.
\item GF Functional types are mapped compositionally : $⟦A → B⟧ = ⟦A⟧ → ⟦B⟧$
\item Every GF syntactic construction function ($f : X$) is mapped to a function $⟦f⟧$ such that $⟦f⟧ : ⟦X⟧$.
\item GF function applications are mapped compositionally: $⟦t(u)⟧ = ⟦t⟧ (⟦u⟧)$.
\end{enumerate}
Because all systems embed the simply-typed lambda calculus, ensuring
type-preservation locally means that types are preserved globally.
Therefore, we are certain that every GF syntax tree can be mapped to
Haskell, and eventuall Coq, without error.

@@

\paragraph{Adjectives}
In our system, we interpret gradable adjectives using a pair of a
measure $m : objects → Z$ and and a threshold $τ : Z$, where $Z$ is
treated as an abstract ordered ring by Coq. (This structure has no
dynamic aspect in our model, and thus is entirely handled within Coq.)
For subsective adjectives, $τ$ will additionally depend on the class
of the object in question. This structure has the benefit that
opposite adjectives can be easily represented (measures are opposites
$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 >
0$). Formalization aside, this idea is reminiscent of degree-based
approaches to gradable adjectives of
\citet{Cresswell:1976,Heim:1985,Kennedy:2007}.  Additionally,
adjectival phrases, as present in the FraCaS suite, are interpreted as
linear inequations in $Z$. Solving systems of such inequations is
decidable. Indeed, the tactic that Coq offers for this purpose can
solve all such problems in the FraCaS suite, automatically.

@@

\paragraph{error analysis}
Our system fails to classify correctly 28 cases out of 259.  We give
here a summary of the missing features which are responsible for the
biggest number of failures. The one biggest source of error is an
incomplete handling of group readings. (FraCaS
013, 014, 046, 084, 111, 124, 126, 127, 137, 171, 172, 191, 193, 195, 243, 250, 333, 346). These
are cases where a syntactic conjunction of individuals is treated as a
semantic group, or where precise counting of the members of a group is
necessary.  Other problematic cases are where a definite plural does
not have a universal reading (091, 094, 095) --- these require an
analysis at the pragmatic level, which our system lacks. Additionally,
neither measure phrases (242) nor attributive comparatives (244, 245)
are handled.

@@

\bibliography{PaperTools/bibtex/jp.bib,phorae.bib,IWCSbib.bib,NLI.bib}
